Thesis Title

### 2022-03-20 chapters/introduction.tex
finish introduction first draft

As explored in \secref{ssec:reads_2_genomes}, longer sequences are more informative than shorter sequencing data, as the one obtained from second-generation sequencing (see \secref{ssec:2nd_gen_seq}), and can provide a more complete picture of the microbial community in a given samples. 
Several dedicated metagenomic assembly tools are available, such as metaSPAdes\footnote{\url{}}  and MegaHIT\footnote{\url{}} \citep{nurk_metaspades_2017, li_megahit_2015}. 
These tools, in comparison to single-cell data assemblers, are better at dealing with the combination of intra and intergenomic repeats and uneven sequencing coverage \citep{olson_metagenomic_2017}.

\section{Trash text}

For virulence gene detection and antimicrobial resistance characterization a mapping approach, with an adequate database, is usually followed, using as reference the Virulence Factors Database (Chen et al., 2016), and ResFinder (Zankari et al., 2012) or CARD (Jia et al., 2017) for antimicrobial resistance gene detection. Besides mapping, other strategies have been applied, like Mash Screen (Ondov et al., 2016), that offer similar results in a faster way.  Similar strategies can be applied to plasmid detection by using the PlasmidFinder (Carattoli et al., 2014) or RefSeq plasmid (O’Leary et al., 2016) databases. The minimap 2 tool (Li, 2018) is a good alternative to map long-read data to any of the resistance and virulence databases mentioned. 
It is possible to genotype the bacterial population in a metagenomic sample, but only for short-read sequencing data.  MetaMLST (Zolfo et al., 2017) reconstructs the MultiLocus Sequence Typing (MLST) loci directly from the sequencing data and provides a pre-built database for the analysis.

\subsection{Metagenome Assembly}

Several limitations arise when using just the sequencing data. Although relatively fast and providing quantitative information, it’s strictly dependent on the content of the databases used. In addition it lacks context information, as linking the characterizing information to a given identified organism isn’t possible.

For long-read data, no dedicated metagenomic assembler is not yet available, but several assembler for long-read data as available, including Canu (Koren et al., 2017) and Unicycler (Wick et al., 2017). The latter allows for hybrid assemblies to be constructed, combining short and long-read information to produce the best assembly possible. Nevertheless, the use of non dedicated assemblers for metagenomics may come with the cost of wrongly interpret variation as error, especially in samples that contained closely related species and the construction of chimeric sequences (Teeling & Glockner, 2012) as traditional assemblers follow the basic principle that the coverage in a sample is constant. 
The assembly-based approach requires the grouping of the different contigs into bins, ideally each collecting the sequences that belong to a microorganism present in the sample. The binning process can be taxonomy dependent, relying on a database to aggregate the sequences, or independent. The independent approach has the benefit of not relying on a database, but instead it uses the composition of each sequence and coverage profiles to cluster together sequences that might belong to the same organism. These algorithms don’t require prior knowledge about the genomes in a given sample, instead relying on features inherent to the sequences in the sample. Although most binning softwares can work with single metagenomic samples, most make use of differential coverage of multiple samples to improve the binning process (Sedlar et al., 2016). It allows the handling of complex ecosystems and might be crucial when analysing samples recovered from sites with a complex microbiota. 
A comparison of five taxonomic independent binning softwares and four taxonomic binning softwares (Sczyrba et al., 2017) revealed that, for taxonomic independent approaches, MaxBin 2.0 (Wu et al., 2016) had the highest completeness and purity in the bins obtained, with ~20\% better results in comparison with the second best ranked tool. For taxonomic binning, working similarly to the direct taxonomic assignment of the sequencing data, PhyloPythiaS+ (Gregor et al., 2016) obtained better results in accuracy, completeness and purity, followed by Kraken (Wood & Salzberg, 2014) that still obtained decent results with the added benefit of very high speed of analysis, ease of use and inclusion of the pre-built databases.
The last step on the assembly methodology is the evaluation of the completeness and contamination of the bins. When using a taxonomic binner, the effects of contamination are mitigated as the sequence clustering is performed based on matches with reference database. The contaminants, if present in the database, will be separated into different bins or just added to the bin of unclassified sequences. When using a taxonomic independent binning software, the composition and abundance might not be enough to discriminate between all the organisms, with the possible result of having bins with contaminating sequences of other organisms present in the sample. CheckM (Parks et al., 2015) assesses the quality of the recovered genomes, estimating completeness and contamination by evaluating ubiquitous single-copy genes. 
Another problem with metagenomic assembly is the high number of ambiguities that fail to being resolved, mostly due to the possible presence of several strains of the same species or species that are closely related. When faced with this ambiguities the assembler usually breaks the sequence, leading to fragmented reconstructions of genomes. MetaQUAST (Mikheenko et al., 2016) that besides computing several metrics to evaluate assembly quality like number of contigs, maximum contig length, etc, also uses reference-based method, either provided by the user or by identifying the appropriate reference sequences by 16S ribosomal RNA identification, to identify mis-assemblies and structural variants. VALET (https://github.com/marbl/VALET) is a de novo pipeline for detecting mis-assemblies without the need for references, relying instead on coverage and length to do the assignment, as well as providing severa visual representations of assembly quality. After the mis-assemblies have been detected, they can be visualized in Icarus (Mikheenko et al., 2016) for metaQUAST, or IGV (interactive genome viewer) for VALET. Anvi’o (Eren et al., 2015) is an analysis and visualization platform that empowers binning refinement and genome completeness and contamination evaluation with interactive interface. 
All downstream processes used in single cell genomes can be applied to each of the resulting binned genomes, that now represent a taxonomic unit recovered from the original metagenomic sample. The typical workflow usually involves antimicrobial resistance and ccvirulence detection. Similar approaches can be used as described in the Direct Taxonomic Assignment and Characterization section by using alignment methods, such as BLAST (Altschul et al., 1990) or DIAMOND (Buchfink et al., 2015), to compare against the Virulence Factors Database, and the ResFinder or CARD databases. Genotyping can be done through the mlst software (https://github.com/tseemann/mlst). The reconstructed genomes allow for the use comparative genomics against other references by using, for example, cgMLST or SNP analysis, and playing a major role in early outbreak detection.

\subsection{Virus in Metagenomic Analysis}
One of the biggest advantages of using metagenomic methods is the detection of not only bacterial organisms, but also viral and eukaryotic pathogens. 
Besides the limitations inherent to the metagenomic process, the retrieval of viral genomes from clinical samples has added difficulties. The fragments of viral genomes are typically orders of magnitude less abundant, the viral genomes often deviate considerably from reference genomes, and the high intrapopulation viral diversity can lead to ambiguous sequence reconstruction or broken assemblies (Rose et al., 2016). Adding to this, the relatively few viral reference genomes can render classification problematic. For fungi, there’s an underrepresentation of the diversity of this group in databases as it remains understudied compared to bacterial microbiomes (Donovan et al., 2018).
Of the tools mentioned for read classification, Kraken’s MiniKraken and MetaPhlAn 2 databases is the most inclusive, including information of virus, bacteria, human and fungi. None other method of the ones described in this review provide a database as inclusive but many, such as Midas, allow the user to build custom databases although requiring very high computational power. Alternatively, the assembly based method can be implemented followed by an alignment search to a database that includes fungi and viral genomes, such as NCBI’s RefSeq (O’Leary et al., 2016) or GenBank (Benson et al., 2005).

>>>>

As explored in \secref{ssec:reads_2_genomes}, longer sequences are more informative than shorter sequencing data, as the one obtained from second-generation sequencing (see \secref{ssec:2nd_gen_seq}), and can provide a more complete picture of the microbial community in a given samples. 
Several dedicated metagenomic assembly tools are available, such as metaSPAdes\footnote{\url{}}  and MegaHIT\footnote{\url{}} \citep{nurk_metaspades_2017, li_megahit_2015}. 
These tools, in comparison to single-cell data assemblers, are better at dealing with the combination of intra and intergenomic repeats and uneven sequencing coverage \citep{olson_metagenomic_2017}.
For third-generation sequencing, dedicated metagenomic assemblers have recently emerged, such as meta-flye\footnote{\url{https://github.com/fenderglass/Flye/}} which expands on the original flye assembler by overcoming a k-mer selection limitation on low abundance species \citep{kolmogorov_metaflye_2020}. Nevertheless, the use of non dedicated assemblers for metagenomics may come with the cost of wrongly interpret variation as error, especially in samples that contained closely related species and the construction of chimeric sequences as traditional assemblers follow the basic principle that the coverage in a sample is constant \citep{teeling_current_2012}.

The assembly-based approach requires the grouping of the different contigs into bins, ideally each collecting the sequences that belong to a microorganism present in the sample. The binning process can be taxonomy dependent, relying on a database to aggregate the sequences, or independent. The independent approach has the benefit of not relying on a database, but instead it uses the composition of each sequence and coverage profiles to cluster together sequences that might belong to the same organism. These algorithms don’t require prior knowledge about the genomes in a given sample, instead relying on features inherent to the sequences in the sample. Although most binning software can work with single metagenomic samples, most make use of differential coverage of multiple samples to improve the binning process \citep{sedlar_bioinformatics_2017}. It allows the handling of complex ecosystems and might be crucial when analysing samples recovered from sites with a complex microbiota. A comparison of five taxonomic independent  and four taxonomic binning software by \cite{sczyrba_critical_2017} revealed that, for taxonomic independent approaches, MaxBin2\footnote{\url{https://sourceforge.net/projects/maxbin2/}} had the highest completeness and purity in the bins obtained \citep{}. For taxonomic binning, working similarly to the direct taxonomic assignment of the sequencing data, PhyloPythiaS+\footnote{\url{https://github.com/algbioi/ppsp}} obtained better results in accuracy, completeness and purity, followed by Kraken\footnote{\url{https://github.com/DerrickWood/kraken2/}} that still obtained decent results with the added benefit of very high speed of analysis, ease of use and inclusion of the pre-built databases \citep{gregor_phylopythias_2016, wood_kraken_2014}.

The last step on the assembly methodology is the evaluation of the completeness and contamination of the bins. When using a taxonomic binner, the effects of contamination are mitigated as the sequence clustering is performed based on matches with reference database. The contaminants, if present in the database, will be separated into different bins or just added to the bin of unclassified sequences. When using a taxonomic independent binning software, the composition and abundance might not be enough to discriminate between all the organisms, with the possible result of having bins with contaminating sequences of other organisms present in the sample. CheckM\footnote{\url{https://ecogenomics.github.io/CheckM/}} assesses the quality of the recovered genomes, estimating completeness and contamination by evaluating ubiquitous single-copy genes \citep{parks_checkm_2015}.

Another problem with metagenomic assembly is the high number of ambiguities that fail to being resolved, mostly due to the possible presence of several strains of the same species or species that are closely related. When faced with this ambiguities the assembler usually breaks the sequence, leading to fragmented reconstructions of genomes. MetaQUAST\footnote{\url{https://github.com/ablab/quast}} that besides computing several metrics to evaluate assembly quality like number of contigs, maximum contig length, etc, also uses reference-based method, either provided by the user or by identifying the appropriate reference sequences by 16S ribosomal RNA identification, to identify mis-assemblies and structural variants \citep{mikheenko_metaquast_2016}. All downstream processes used in whole genome sequencing draft genomes can be applied to each of the resulting binned genomes, that now represent a taxonomic unit recovered from the original metagenomic sample.
